Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-20 paper/paper.tex
Changes to sections 5.1 and 5.2

To evaluate our algorithms, we developed a cloud workflows simulator based on
CloudSim~\cite{Calheiros11}. To represent our infrastructure model, we defined
Cloud, VM and WorkflowEngine simulation entities. Cloud object is responsible
for staring and terminating VMs using API similar to Amazon EC2. VM object is
responsible for execution of individual tasks, while WorkflowEngine deals with
management of ensembles and workflows. WorkflowEngine uses separate, although
tightly coupled Scheduler and Provisioner modules. We assume that a single core
of a VM executed submitted tasks sequentially and that a task is given exclusive
access to allocated VM. Although CloudSim provides a more advanced
infrastructure model which includes time-sharing and space-sharing policies and
power consumption in a datacenter, we do not use these features, since we are
interested mainly in execution of tasks on VMs and high-level workflow scheduling
on top of that. The simulator relies on DAG and DAX format used by Pegasus
project and workflow gallery.

@@

We have prepared test scenarios to evaluate the proposed algorithms under
such a range of parameters that allows to observe the interesting properties of
algorithms for smaller and larger constraints. For a given ensemble we can
expect in general that the number of work done (in terms of workflows completed)
mainly depends on the budget constraint, which determines the number of
machine-hours that can be provisioned. The ideal but trivial case is then a
deadline long enough to allow running all the workflow tasks sequentially: this
eliminates all the overheads of parallel execution and leads to the highest
resource utilization. More interesting are the cases when the deadline is
shorter, since it forces provisioning of multiple VMs and thus increases
overheads due to parallel execution so the budget cannot be spent so
efficiently.

@@

We assumed that the interesting number of workflows in ensemble should be in the
order of between 10 and 100 workflows. This is motivated by several reasons:
first, such numbers are typical for the real applications, e.g. the number of geographical sites for
CyberShake workflow is in the order of 100. Second, the smaller ensembles
consisting of just a few workflows can be aggregated in a single workflow so
there is no need to treat them as ensembles. Similarly, when the number of
workflows is larger that 100 and each workflow has a large number of parallel
tasks, the problem of efficiently allocating them to the resources becomes much
easier and similar to the bag-of-tasks problem.

@@

In order to evaluate the algorithms on a standard set of workflows, we used
ensembles constructed from the DAGs available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
in the workflow generator gallery~\cite{Bharathi08}. The gallery contains
workflows of five applications from Pegasus as well as several other applications 
from~\cite{Ramakrishnan08}. We constructed ensembles of 100 DAGs by sampling the
workflows of different sizes based on a given distribution. By the size of the
workflow means the number of jobs. Two distributions are considered:
\begin{itemize}
  \item ensemble of workflows of equal size (constant)
  \item Pareto-like distribution.
\end{itemize}

Pareto-like distribution gives a small number of workflows with large size and a
large number of workflows with small size, as seen in
Fig.~\ref{fig:ensemble-distribution}. The number of largest
workflows (of size $\geq$ 900) is slightly increased to represent the
``long-tail'', i.e. that there is a non-insignificant number of large workflows
in the ensemble. The workflows are sorted according to their size and the largest workflows are
given the highest priority.

@@

In the tests we we assumed that all the VMs are homogeneous and have the
processing speed of 1000 MIPS (million instructions per second), the price for
one VM-hour is \$1 and that the sizes of tasks in the workflow gallery are given
in seconds measured on 1000 MIPS machines. In this study we do not take into
account the heterogeneity of the infrastructure since we assume that it is
always possible to select a VM type that has the best price to performance ratio
for a given application.

For each application type, we selected such ranges of constraints (deadline,
budget) that it is possible to observe the possibly broad result space: from the
tight constraints, when only a small number of workflows can finish to the
liberal constraints when all or almost all the workflows can finish. These
parameters vary among the applications. Montage and CyberShake benchmarks
workflows required the budget in \$20 -- \$150 range and deadlines from 1 to
20 hours to complete the ensemble. LIGO and SIPHT workflows have more long
tasks, so they require budgets up to \$1,500 and deadlines extended to 40
hours. The Epigenomics workflow is even more costly and needs approximately \$12,000
and 1500 hours to complete the full ensemble.

We show the results for three algorithms: DPDS, WA-DPDS and
SPSS. These tests were run with maximum autoscaling factor set to 1.0 for
DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that due to
high parallelism of workflows the resource utilization remains high most of the time, 
which leads to rapid provisioning of new resources at the beginning of ensemble
execution. This in turn results in less efficient total utilization due to high
level of parallelism. Setting the scaling limit to 1.0, ensures that the shape
of resource provisioning plan is closest to a rectangle (see
Fig.~\ref{fig:spds-example}) and the VM termination procedure
(Algorithm~\ref{alg:prov}) allows maximum utilization of VMs before shutdown.

>>>>

To evaluate and compare our algorithms, we developed a cloud workflow simulator based on
CloudSim~\cite{Calheiros11}. Our simulation model consists of Cloud, VM and WorkflowEngine 
entities. The Cloud entity is responsible for staring and terminating VM entities using 
an API similar to Amazon EC2. VM entities simulate the execution of individual tasks,
and the WorkflowEngine entity manages the scheduling of tasks and the provisioning of
VMs. We assume that the VMs have a single core and execute tasks sequentially. Although 
CloudSim provides a more advanced infrastructure model which includes time-sharing and 
space-sharing policies, and power consumption in a datacenter, we do not use these 
features, since we are interested mainly in the execution of tasks on VMs and high-level 
workflow scheduling and provisioning. The simulator reads workflows in a modified version of
the DAX format used by the Pegasus Workflow Management System~\cite{Deelman2005} that was
created for the synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi08}.

@@

We have prepared test scenarios to evaluate the proposed algorithms under a 
range of parameters that enables us to observe the interesting characteristics 
of the different algorithms for smaller and larger constraints. For a given 
ensemble we can expect, in general, that the amount of work done (in terms 
of workflows completed) depends mainly on the budget constraint, which 
determines the number of machine-hours that can be provisioned. The deadline, 
although important in terms of scheduling, is assumed to be greater than the 
critical path of every workflow. The ideal but trivial case, then, is a deadline 
long enough to allow all the workflow tasks to be executed sequentially: this 
eliminates all the overheads of parallel execution and leads to the highest 
resource utilization. More interesting, however, are the cases where the deadline 
is shorter than the serial schedule, since it forces the system to provision 
multiple VMs to execute workflows in parallel.

@@

The number of workflows in an ensemble depends on the user's requirements, but
we assume that ensemble sizes are on the order of between 10 and 100 workflows.
This range is motivated by several reasons: First, such sizes are typical of the
real applications we have examined. For example, the number of geographical 
sites of interest to the users of the CyberShake application is on the order 
of 100. Second, smaller ensembles consisting of just a few workflows can be 
aggregated into a single workflow, so there is no need to treat them as an
ensemble. Similarly, when the number of workflows is larger than 100, and each 
workflow has a large number of tasks, either the constraints are low enough to
prevent many of the workflows from running, or the problem of efficiently 
allocating them to the resources becomes similar to the bag-of-tasks problem, 
which is easier to solve efficiently.

@@

In order to evaluate the algorithms on a standard set of workflows, we used
ensembles constructed from the workflows available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
in the workflow generator gallery~\cite{Bharathi08}. The gallery contains synthetic
workflows modeled on real applications. We used workflows from five applications 
that were developed for the Pegasus Workflow Management System. We constructed 
ensembles of 100 workflows by sampling collections of workflows of various sizes
(number of tasks) based on a given distribution. Two distributions are considered:

\begin{itemize}
  \item constant distribution
  \item Pareto-like distribution.
\end{itemize}

The constant distribution produces ensembles containing of workflows of equal size, 
while the Pareto-like distribution produces ensembles with a small number of larger 
workflows and a large number of smaller workflows, as seen in 
Fig.~\ref{fig:ensemble-distribution}. The number of large workflows (of size $\geq$ 900) 
is slightly increased to represent the ``heavy-tail'', i.e. that there is a 
not-insignificant number of large workflows in the ensemble. We assume that large
workflows are more important to users, so in the Pareto ensembles the workflows are 
assigned priorities according to their size so that the largest workflows are given 
the highest priority. In future work we plan to explore other ensemble types such as
uniform distributions.

@@

In the experiments we assumed that all the VMs are homogeneous, have a
processing speed of 1000 MIPS (million instructions per second), and a price of
\$1 per VM-hour. We also assume that the sizes of tasks in the workflow gallery 
are given in seconds measured on a 1000 MIPS machine. In this study we do not 
take into account the heterogeneity of the infrastructure since we assume that 
it is always possible to select a VM type that has the best price to performance 
ratio for a given application.

For each application, we selected ranges of constraints (deadline and budget) so
that it is possible to observe the possibly broad result space: from tight 
constraints, where only a small number of workflows can be completed, to 
more liberal constraints where all, or almost all, of the workflows can be 
completed. These parameters vary among the applications. Montage and CyberShake 
ensembles required budgets in the \$20 -- \$150 range and deadlines from 1 to
20 hours to be completed. The LIGO and SIPHT workflows have more long-running
tasks, so their ensembles required budgets up to \$1,500 and deadlines up to 
40 and 60 hours. The Epigenomics workflow is even more costly, requiring 
approximately \$12,000 and 1500 hours to complete a full ensemble.

We show the results for three algorithms: DPDS, WA-DPDS and SPSS. These 
experiments were run with maximum autoscaling factor set to 1.0 for
DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that, 
due to the high parallelism of workflows, the resource utilization remains 
high most of the time, which leads to the rapid provisioning of new resources 
at the beginning of ensemble execution. This, in turn, results in less 
efficient total utilization due to high levels of parallelism. Setting the 
scaling limit to 1.0, ensures that the shape of the resource provisioning 
plan  is close to a rectangle (see Fig.~\ref{fig:spds-example}) and the VM 
termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
of VMs before shutdown.
